\vspace{-10pt}
\section{Preliminaries}
\label{sec:background}
\vspace{-8pt}
The goal in reinforcement learning is to learn a policy that maximizes the expected cumulative discounted reward in a Markov decision process (MDP), which is defined by a tuple $(\mathcal{S}, \mathcal{A}, \transitions, r, \gamma)$.
$\mathcal{S}, \mathcal{A}$ represent state and action spaces, $\transitions(\bs' | \bs, \ba)$ and $r(\bs,\ba)$ represent the dynamics and reward function, and $\gamma \in (0,1)$ represents the discount factor. $\behavior(\ba|\bs)$ represents the behavior policy, $\mathcal{D}$ is the dataset, and $d^\behavior(\bs)$ is the discounted marginal state-distribution of $\behavior(\ba|\bs)$. The dataset $\mathcal{D}$ is sampled from $d^\behavior(\bs) \behavior(\ba|\bs)$. {On all states $\bs \in \mathcal{D}$, let $\hatbehavior(\ba|\bs_) := \frac{\sum_{\bs,\ba \in \mathcal{D}} \mathbf{1} [\bs = \bs , \ba = \ba]}{\sum_{\bs \in \mathcal{D}} \mathbf{1}[\bs = \bs]}$ denote the empirical behavior policy, at that state.} We assume that the rewards $r$ satisfy: $|r(\bs, \ba)| \leq R_{\max}$.

Off-policy RL algorithms based on dynamic programming maintain a parametric Q-function $Q_\theta(s, a)$ and, optionally, a parametric policy, $\pi_\phi(a|s)$. 
Q-learning methods train the Q-function by iteratively applying the Bellman optimality operator \mbox{$\mathcal{B}^*Q(\bs, \ba) = r(\bs, \ba) + \gamma \E_{\bs' \sim P(\bs'|\bs, \ba)}[\max_{\ba'} Q(\bs', \ba')]$}, and use exact or an approximate maximization scheme, such as CEM~\citep{kalashnikov2018qtopt} to recover the greedy policy. In an actor-critic algorithm, a separate policy is trained to maximize the Q-value.
Actor-critic methods alternate between computing $Q^\policy$ via (partial) policy evaluation,
by iterating the Bellman operator, $\mathcal{B}^\pi Q= r + \gamma P^\pi Q$, where $P^\pi$ is the transition matrix coupled with the policy: $P^\pi Q(\bs, \ba) = \E_{\bs' \sim \transitions(\bs' | \bs, \ba), \ba' \sim \pi(\ba'|\bs')} \left[ Q(\bs', \ba') \right],$
and improving the policy
$\policy(\ba|\bs)$ by updating it towards actions that maximize the expected Q-value. Since $\mathcal{D}$ typically does not contain all possible transitions $(\bs, \ba, \bs')$, the policy evaluation step actually uses an empirical Bellman operator that only backs up a single sample. We denote this operator $\hat{\bellman}^\policy$. Given a dataset $\mathcal{D} = \{(\bs, \ba, r \bs')\}$ of tuples from trajectories collected under a behavior policy $\behavior$:
\begin{small}
\begin{align*}
    \hat{Q}^{k+1} \leftarrow& \arg\min_{Q} \E_{\bs, \ba, \bs' \sim \mathcal{D}}\left[ \left((r(\bs, \ba) + \gamma \E_{\ba' \sim \hat{\policy}^k(\ba'|\bs')}[\hat{Q}^{k}(\bs', \ba')]) - Q(\bs, \ba)\right)^2 \right]~\text{(policy evaluation)}\\ 
    \hat{\policy}^{k+1} \leftarrow& \arg\max_{\policy} \E_{\bs \sim \mathcal{D}, \ba \sim \policy^k(\ba|\bs)}\left[\hat{Q}^{k+1}(\bs, \ba)\right]~~~ \text{(policy improvement)} 
\end{align*}
\end{small}
Offline RL algorithms based on this basic recipe suffer from action distribution shift~\citep{kumar2019stabilizing,wu2019behavior,jaques2019way,levine2020offline} during training, because the target values for Bellman backups in policy evaluation use actions sampled from the learned policy, $\policy^k$, but the Q-function is trained only on actions sampled from the behavior policy that produced the dataset $\mathcal{D}$, $\behavior$. Since $\policy$ is trained to maximize Q-values, it may be biased towards out-of-distribution (OOD) actions with erroneously high Q-values. In standard RL, such errors can be corrected by attempting an action in the environment and observing its actual value. However, the inability to interact with the environment makes it challenging to deal with Q-values for OOD actions in offline RL. Typical offline RL methods~\citep{kumar2019stabilizing,jaques2019way,wu2019behavior,siegel2020keep} mitigate this problem by constraining the learned policy~\citep{levine2020offline} away from OOD actions. {Note that Q-function training in offline RL does not suffer from state distribution shift, as the Bellman backup never queries the Q-function on out-of-distribution states. However, the policy may suffer from state distribution shift at test time.}
