
\section{Preliminaries}

\subsection{Offline reinforcement learning}
We consider learning in a Markov decision process (MDP) described by the tuple ($\mathcal{S}$, $\mathcal{A}$, $P$, $\mathcal{R}$). The MDP tuple consists of states $s \in \mathcal{S}$, actions $a \in \mathcal{A}$, transition dynamics $P(s'|s,a)$, and a reward function $r = \mathcal{R}(s,a)$. 
We use $s_t$, $a_t$, and $r_t=\mathcal{R}(s_t, a_t)$ to denote the state, action, and reward at timestep $t$, respectively. A trajectory is made up of a sequence of states, actions, and rewards: $\tau = (s_0, a_0, r_0, s_1, a_1, r_1, \ldots, s_T, a_T, r_T)$. The return of a trajectory at timestep $t$, $R_t = \sum_{t'=t}^T r_{t'}$, is the sum of future rewards from that timestep.
The goal in reinforcement learning is to learn a policy which maximizes the expected return $\mathbb{E}\Bigl[\sum_{t=1}^T r_t\Bigr]$ in an MDP.
In offline reinforcement learning, instead of obtaining data via environment interactions, we only have access to some fixed limited dataset consisting of trajectory rollouts of arbitrary policies.
This setting is harder as it removes the ability for agents to explore the environment and collect additional feedback.

\subsection{Transformers}
Transformers were proposed by \citet{vaswani2017attention} as an architecture to efficiently model sequential data.
These models consist of stacked self-attention layers with residual connections.
Each self-attention layer receives $n$ embeddings $\{x_i\}_{i=1}^n$ corresponding to unique input tokens, and outputs $n$ embeddings $\{z_i\}_{i=1}^n$, preserving the input dimensions.
The $i$-th token is mapped via linear transformations to a key $k_i$, query $q_i$, and value $v_i$.
The $i$-th output of the self-attention layer is given by weighting the values $v_j$ by the normalized dot product between the query $q_i$ and other keys $k_j$:
\begin{equation}
    z_i = \sum_{j=1}^n \texttt{softmax}(\{\langle q_i, k_{j'} \rangle\}_{j'=1}^n)_j \cdot v_j.
\end{equation}
As we shall see later, this allows the layer to assign ``credit'' by implicitly forming state-return associations via similarity of the query and key vectors (maximizing the dot product).
In this work, we use the GPT architecture \citep{radford2018gpt}, which modifies the transformer architecture with a causal self-attention mask to enable autoregressive generation, replacing the summation/softmax over the $n$ tokens with only the previous tokens in the sequence ($j \in [1, i]$).
We defer the other architecture details to the original papers.
