\documentclass{article}

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{amsmath}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{property}{Property}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\newcommand{\E}{\mathbb{E}}
\newcommand{\Ls}{\mathcal{L}}
\newcommand{\gN}{\mathcal{N}}
\newcommand{\calO}{\mathcal{O}}
\newcommand{\R}{\mathbb{R}}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\def\[#1\]{\begin{align}#1\end{align}}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

\usepackage[accepted]{icml2019}

\icmltitlerunning{Set Transformer}

\begin{document}

\twocolumn[
\icmltitle{Set Transformer: A Framework for Attention-based \\Permutation-Invariant Neural Networks}

\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Juho Lee}{oxford1,aitrics}
\icmlauthor{Yoonho Lee}{kakao}
\icmlauthor{Jungtaek Kim}{postech}
\icmlauthor{Adam R. Kosiorek}{oxford1,oxford2}
\icmlauthor{Seungjin Choi}{postech}
\icmlauthor{Yee Whye Teh}{oxford1}
\end{icmlauthorlist}

\icmlaffiliation{oxford1}{Department of Statistics, University of Oxford, United Kingdom}
\icmlaffiliation{oxford2}{Oxford Robotics Institute, University of Oxford, United Kingdom}
%\icmlaffiliation{oxford2}{Applied Artificial Intelligence Lab, Oxford Robotics Institute, University of Oxford}
\icmlaffiliation{postech}{Department of Computer Science and Engineering, POSTECH, Republic of Korea}
\icmlaffiliation{aitrics}{AITRICS, Republic of Korea}
\icmlaffiliation{kakao}{Kakao Corporation, Republic of Korea}

\icmlcorrespondingauthor{Juho Lee}{juho.lee@stats.ox.ac.uk}

\icmlkeywords{Set transformer, Set-taking network, Attention-based network, Permutation-invariant network}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
Many machine learning tasks such as multiple instance learning, 3D shape recognition and few-shot image classification are defined on sets of instances.
Since solutions to such problems do not depend on the order of elements of the set, models used to address them should be \textit{permutation invariant}.
We present an attention-based neural network module, the \textit{Set Transformer}, specifically designed to model interactions among elements in the input set.
The model consists of an encoder and a decoder, both of which rely on attention mechanisms.
In an effort to reduce computational complexity, we introduce an attention scheme inspired by inducing point methods from sparse Gaussian process literature.
It reduces computation time of self-attention from quadratic to linear in the number of elements in the set.
We show that our model is theoretically attractive and we evaluate it on a range of tasks, demonstrating increased performance compared to recent methods for set-structured data.
\end{abstract}

\input{01_introduction.tex}
\input{02_background.tex}
\input{03_main.tex}
\input{04_related_works.tex}
\input{05_experiments.tex}
\input{06_conclusion.tex}


\paragraph{Acknowledgments}
JL and YWT's research leading to these results has received funding from the
European Research Council under the European Union's Seventh Framework
Programme (FP7/2007-2013) ERC grant agreement no. 617071.
JL has also received funding from EPSRC under grant EP/P026753/1. JL acknowledges support from IITP grant funded by the Korea government(MSIT) (No.2017-0-01779, XAI) and Samsung Research Funding \& Incubation Center of Samsung Electronics under Project Number SRFC-IT1702-15.

\bibliography{set_transformer}
\bibliographystyle{icml2019}

\end{document}

%\appendix
%\input{10_appendix.tex}
