\begin{abstract}

Advances in 3D reconstruction and novel view synthesis have enabled efficient and photorealistic rendering.  
However, images for reconstruction are still either largely manual or constrained by simple preplanned trajectories.
To address this issue, recent works propose generalizable next-best-view planners that do not require online learning.
Nevertheless, robustness and performance remain limited across various shapes. 
Hence, this study introduces Voxel-Face-Aware \underline{H}ierarchical N\underline{e}xt-Be\underline{s}\underline{t}-V\underline{i}ew \underline{A}cquisition for Efficient 3D Reconstruction (Hestia\footnote{In Greek mythology, Hestia is the goddess of the hearth, symbolizing home, foundation, and structure, representing stability and guidance in complex systems}), which addresses the shortcomings of the reinforcement learning-based generalizable approaches for five-degree-of-freedom viewpoint prediction. 
Hestia systematically improves the planners through four components: a more diverse dataset to promote robustness, a hierarchical structure to manage the high-dimensional continuous action search space, a close-greedy strategy to mitigate spurious correlations, and a face-aware design to avoid overlooking geometry.
Experimental results show that Hestia achieves non-marginal improvements, with at least a 4\% gain in coverage ratio, while reducing Chamfer Distance by 50\% and maintaining real-time inference. 
In addition, Hestia outperforms prior methods by at least 12\% in coverage ratio with a 5-image budget and remains robust to object placement variations. 
Finally, we demonstrate that Hestia, as a next-best-view planner, is feasible for the real-world application.
%The code and data processing script for Hestia are provided in the supplementary materials and will be released after publication.
Our project page is \href{https://johnnylu305.github.io/hestia_web}{\texttt{https://johnnylu305.github.io/hestia\_web}}.
%\noindent\textbf{Project page:} %\href{https://github.com/johnnylu305/Hestia-NBV}%{\texttt{https://github.com/johnnylu305/Hestia-NBV}}
\end{abstract}
