\pdfoutput=1

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2022

\PassOptionsToPackage{numbers, square}{natbib}

% ready for submission
\usepackage[final]{neurips_2022}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{bbold}
\usepackage{subfig}
\usepackage{color}
\usepackage{enumitem}
\usepackage{comment}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2022}


% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2022}


% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2022}


\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{dsfont}         % for \IR, etc.
\usepackage{xspace}

\usepackage{amsthm}

\setcitestyle{numbers}

\newtheorem{theorem}{Theorem}
\newtheorem*{T1}{Theorem~\ref{thm:volumetric_approximation_of_coverage}}
\newtheorem{lemma}[theorem]{Lemma}


\title{\method: Surface Coverage Optimization in Unknown Environments by Volumetric Integration}

% \title{\method: Surface Coverage Optimization in Unknown Environments by Integrating over a Volumetric Representation}

% \title{\method: Surface Coverage Optimization in Unknown Environments with Implicit Monte Carlo Integration}


% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.


\author{%
  Antoine Guédon \qquad Pascal Monasse \qquad Vincent Lepetit %\thanks{Use footnote for providing further information about author (webpage, alternative address)---\emph{not} for acknowledging funding agencies.} 
  \\
  % \quad \\
  LIGM, Ecole des Ponts, Univ Gustave Eiffel, CNRS, France \\
  \texttt{\{antoine.guedon,pascal.monasse,vincent.lepetit\}@enpc.fr} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\input{commands}

\addeditor{vincent}{VL}{0.0, 0.5, 0.0}
\addeditor{antoine}{AG}{0.0, 0.0, 0.5}
\showeditsfalse

\newcommand{\method}{SCONE\xspace}
\newcommand{\occfield}{\sigma}
\newcommand{\probfield}{\hat{\sigma}} % Is the notation \sigma OK for the occupancy probability field?
\newcommand{\absvis}{v} % \absvis is the "absolute" visibility
\newcommand{\vis}{\nu} % and \vis is actually the visibility gain
\newcommand{\newvis}{g}  % {\overline{\nu}}
\newcommand{\predvis}{\hat{\nu}}
\newcommand{\camhist}{H}
\newcommand{\proxy}{\hat{\chi}}

\newcommand{\softmax}{\text{softmax}}
\newcommand{\e}{\text{e}}

\textvars{pos,rot}

% Limit is now 10 pages!!!!!!!!

\begin{document}

\maketitle

\input{0b_abstract}

\input{camera_ready_1_intro}

\input{camera_ready_2_approach}

\input{camera_ready_3_experiments}

\input{camera_ready_5_conclusion}

\section*{Acknowledgements}

This work was granted access to the HPC resources of IDRIS under the allocation 2022-AD011013387 made by GENCI. We thank Renaud Marlet, Elliot Vincent, Romain Loiseau, and Tom Monnier for inspiring discussions and valuable feedback.

\bibliographystyle{plainnat}
\bibliography{camera_ready}

%\begin{comment}
\section*{Checklist}


% %%% BEGIN INSTRUCTIONS %%%
% The checklist follows the references.  Please
% read the checklist guidelines carefully for information on how to answer these
% questions.  For each question, change the default \answerTODO{} to \answerYes{},
% \answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
% justification to your answer}, either by referencing the appropriate section of
% your paper or providing a brief inline description.  For example:
% \begin{itemize}
%   \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
%   \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
%   \item Did you include the license to the code and datasets? \answerNA{}
% \end{itemize}
% Please do not modify the questions and only use the provided macros for your
% answers.  Note that the Checklist section does not count towards the page
% limit.  In your paper, please delete this instructions block and only keep the
% Checklist section heading above along with the questions/answers below.
% %%% END INSTRUCTIONS %%%


\begin{enumerate}


\item For all authors...
\begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \answerYes{}{}
  \item Did you describe the limitations of your work?
    \answerYes{} See the last section.
  \item Did you discuss any potential negative societal impacts of your work?
    \answerYes{} A discussion can be found in the supplementary material.
  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
    \answerYes{}
\end{enumerate}


\item If you are including theoretical results...
\begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    \answerYes{}{}
        \item Did you include complete proofs of all theoretical results?
    \answerYes{}
\end{enumerate}


\item If you ran experiments...
\begin{enumerate}
  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    \answerNo{} Not yet, but we will release our source code and dataset.
  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    \answerYes{} In the supplementary material.
        \item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    \answerNo{}
        \item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    \answerYes{} See supplementary material.
\end{enumerate}


\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
\begin{enumerate}
  \item If your work uses existing assets, did you cite the creators?
    \answerYes{}
  \item Did you mention the license of the assets?
    \answerYes{}
  \item Did you include any new assets either in the supplemental material or as a URL?
    \answerNo{}
  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    \answerYes{}
  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    \answerNo{} Our data do not have such issue.
\end{enumerate}


\item If you used crowdsourcing or conducted research with human subjects...
\begin{enumerate}
  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
    \answerNA{}
  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
    \answerNA{}
  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
    \answerNA{}
\end{enumerate}

\end{enumerate}
%\end{comment}

\input{camera_ready_6_appendix}

\end{document}
